Preparing for Potential Closure of European Airspaces due to Re-entering Space Objects

### 2025-06-08 paper.qmd
address reviewer 4 comment on possibility to use transponder data by adding details about NM trajectories and shortcomings of ADS-B data, namely coverage and static aircraft at airports (inflating density).

format:
  ieee-pdf:
    keep-tex: true
    colorlinks: false
    # fig-pos: 'H'
    journaltype: conference
    papersize: a4
    include-in-header:
      - file: media/preamble.tex
    include-before-body:
      - text: |
          \let\qty\SI
          % NOTE: make siunitx's \qty work without clashing with 
          %      the one in physics
  ieee-html: 
    html-math-method: mathjax
    include-in-header:
      file: media/siunitx.html
author:
  - id: es
    name: Enrico Spinielli
    affiliations:
      - ref: aiu
    orcid: 0000-0001-8584-9131
    email: enrico.spinielli@eurocontrol.int
    attributes:
      corresponding: true
    role: 
      - conceptualization
      - writing
      - software
  - name: Rainer Koelle
    affiliations:
      - ref: aiu
    orcid: 0000-0003-2810-1404
    email: rainer.koelle@eurocontrol.int
    role: 
      - editing
  - name: Quinten Goens
    affiliations:
      - ref: aiu
    orcid: 0000-0003-2187-1924
    email: quinten.goens@eurocontrol.int
    role: 
      - editing
affiliations:
  - id: aiu
    name: EUROCONTROL
    department: Aviation Intelligence Unit
    city: Brussels
    country: Belgium
  - id: fake
    name: ACME
    department: Best Tricks
    city: Los Angeles
    country: USA
bibliography: media/references.bib
acronyms:
  insert_loa: false
  fromfile:
    - ./media/acronyms.yml
editor: 
  markdown: 
    wrap: 72
---
```{r}
```

@@

[^adsb]: Crowdsourced ADS-B/Mode S data suffer from lack of coverage away from landmasses and from the need to filter away position reports for stationary aircraft at airports which unnecessarily inflate the density of flying traffic.

This work focusses on the European region. The hourly flight density
estimation is based on flown \acr{IFR} trajectories as recorded by
EUROCONTROL's \acr{NM}[^1]. To decrease spatial complexity, the data is
binned using Uber's H3 geospatial indexing system [@uber-h3]. The
reduction of the spatial complexity enables fast processing. For this
work, we employ a H3-cell-resolution of 3. This corresponds to hexagons
of an average area $\qty{12393}{km^2}$ and an average edge length of
$\qty{68.979}{km}$ [@h3resolutions] as depicted in @fig-cells.

>>>>

format:
  ieee-pdf:
    keep-tex: true
    colorlinks: false
    # fig-pos: 'H'
    journaltype: conference
    papersize: a4
    include-in-header:
      - file: media/preamble.tex
    include-before-body:
      - text: |
          \let\qty\SI
          % NOTE: make siunitx's \qty work without clashing with 
          %      the one in physics
  ieee-html: 
    html-math-method: mathjax
    include-in-header:
      file: media/siunitx.html
author:
  - id: es
    name: Enrico Spinielli
    affiliations:
      - ref: aiu
    orcid: 0000-0001-8584-9131
    email: enrico.spinielli@eurocontrol.int
    attributes:
      corresponding: true
    role: 
      - conceptualization
      - writing
      - software
  - name: Rainer Koelle
    affiliations:
      - ref: aiu
    orcid: 0000-0003-2810-1404
    email: rainer.koelle@eurocontrol.int
    role: 
      - editing
  - name: Quinten Goens
    affiliations:
      - ref: aiu
    orcid: 0000-0003-2187-1924
    email: quinten.goens@eurocontrol.int
    role: 
      - editing
affiliations:
  - id: aiu
    name: EUROCONTROL
    department: Aviation Intelligence Unit
    city: Brussels
    country: Belgium
  - id: fake
    name: ACME
    department: Best Tricks
    city: Los Angeles
    country: USA
  - id: empty

bibliography: media/references.bib
acronyms:
  insert_loa: false
  fromfile:
    - ./media/acronyms.yml
editor: 
  markdown: 
    wrap: 72
---
```{r}
```

@@

[^adsb]: Crowdsourced ADS-B/Mode S data suffer from lack of coverage away from landmasses and from the need to filter away position reports for stationary aircraft at airports which unnecessarily inflate the density at the location.

This work focusses on the European region. The hourly flight density
estimation is based on _flown_ \acr{IFR} trajectories as recorded
by EUROCONTROL's \acr{NM}[^1]. These trajectories are reconstructed from
the filed \acr{FP} and recalculated when big deviations are detected from
position reports received from \acr{ATC}.

To decrease spatial complexity, the data is
binned using Uber's H3 geospatial indexing system [@uber-h3]. The
reduction of the spatial complexity enables fast processing. For this
work, we employ a H3-cell-resolution of 3. This corresponds to hexagons
of an average area $\qty{12393}{km^2}$ and an average edge length of
$\qty{68.979}{km}$ [@h3resolutions] as depicted in @fig-cells.
